Bayes network classifiers for fMRI

### 2012-11-08 article/article.tex
ins alinea line

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.
%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}
\label{sec:introduction}
\maketitle
%% Now the actual contents of first section
Human functional brain mapping is a field of study that has grown strongly in the past decades. This is not a surprise, because it is one of the few ways to look at brain activity of a living animal. There are several techniques available to scan for brain activity in a living human. Of these techniques, functional magnetic resonance imaging (fMRI) has become more and more popular because it has a very good spacial resolution (contrary to MEG/EEG), and it is non-invasive (contrary to PET). Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine. This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.\\
\indent The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds. However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.\\
\indent Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance. Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}. Experimental conditions are the independent variables in this setup, and a general linear model is fitted to each voxel, or group of voxels. By using this technique voxels or brain areas can be identified that differ in average BOLD response across conditions.\\
\indent Alternatively, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition. Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns. When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
\indent These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}. The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline. Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}. \\

>>>>

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.
%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}
\label{sec:introduction}
\maketitle
%% Now the actual contents of first section
Human functional brain mapping is a field of study that has grown strongly in the past decades. This is not a surprise, because it is one of the few ways to look at brain activity of a living animal. There are several techniques available to scan for brain activity in a living human. Of these techniques, functional magnetic resonance imaging (fMRI) has become more and more popular because it has a very good spatial resolution (contrary to MEG/EEG), and it is non-invasive (contrary to PET). Brain activity is inferred from differences in the ratio between oxygenated and deoxygenated blood; when brain activity occurs, more oxygenated blood will flow to a brain region, and this change is detected by the machine. This is called the blood oxygen level dependant (BOLD) response \cite{sm:2012fMRI}.\\
\indent The fMRI scanner measures the activity in three-dimensional pixels called voxels, which are about 3 x 3 x 3 mm in size, and a brain scan is made somewhere between every 500ms and every few seconds. However, before analysing the data is it usually pre-processed to account for head movements by participants in the scanner, and to make between-subjects comparisons possible \cite{sl:2009rl}.\\
\indent Traditionally, data from each voxel, or from a group of voxels, are analysed with analysis of variance. Voxels are usually convoluted with a haemodynamic response function in an attempt to infer the spiking neural activity from the slowly building and dropping BOLD response \cite{he:2002na}. Experimental conditions are the independent variables in this setup, and a general linear model is fitted to each voxel, or group of voxels. By using this technique voxels or brain areas can be identified that differ in average BOLD response across conditions.\\
\indent Insert Alinea about large scale fMRI studies indicating that "everything is significant", thus identifying patterns makes more sense than identifying areas that differ from baseline. \\
\indent Alternatively, Machine learning algorithms have been employed to learn the patterns of activity associated with some experimental condition. Subsequently, these learned patterns can be used on unseen subjects in order to predict experimental conditions from their brain activity patterns. When these predictions are accurate then it means that the identified patterns hold information about the brain activity associated with some experimental task.\\
\indent These machine learning techniques have been successfully used for decoding the orientation of a stimulus from fMRI scans \cite{ka:2005decoding}, identifying different patterns of activity for animate versus inanimate objects \cite{kr:2008RSA}, and even reconstructing visual input based on information in the fMRI scans \cite{mi:2008re}. The reason that these machine learning techniques work so well is that patterns of voxel activity can be identified rather than a significant deviation from some baseline. Specifically, information can be present in the pattern of voxels that belong to a certain brain region, even when the average signal from that region is zero (i.e., on average it does not differ from the baseline condition) \cite{to:2012rev}. \\
